### 2021-01-26 README.Rmd
small readme edits
mostly spellcheck, add library(tidyr)

This package is designed to work for spatially mapped, repeat censused forests plots. The package has commands to fit models of tree growth based on neighborhood competition which can be used to estimate species-specific competition coefficients. The models fits can then be evaluated using a spatial cross-validation scheme to detect possible overfitting. Additionally these models can test whether the species identity of competitors matters using a permutation test-style shuffling of competitor identity (under the null hypothesis) and subsequently evaluating if model performance changes. See Allen and Kim (2020) [A permutation test and spatial cross-validation approach to assess models of interspecific competition between trees](https://doi.org/10.1371/journal.pone.0229930) for a full description.

@@

```{r}
library(tidyverse)
library(forestecology)
library(sf)
library(sfheaders)
library(blockCV)
library(yardstick)
library(snakecase)
library(patchwork)
```

@@

The starting point of our analysis are data from two repeat censuses `census_1_ex` and `census_2_ex`. For example consider the forest census data in `census_1_ex`

@@

In the visualization below, the spatial blocks that act as our cross-validation folds are delineated in orange. Correspondingly the shape of each point indicates which fold each tree has been assigned to.

@@

Based on our `growth` data frame, we now explicitly define all "focal" trees and their respective "competitor" trees in a `focal_vs_comp` data frame. This data frame has rows corresponding to each focal tree and all information about its competitors are saved in list-column variable `comp`. We implemented this nested format using `tidyr::nest()` in order to minimize redundancy, given that the same tree can act as a competitor multiple times.

@@

Using `unnest()` we can fully expand the competitor information saved in the `focal_vs_comp` data frame. For example the tree with `focal_ID` equal to 2 located at (1.5, 2.5) has two competitors within `comp_dist` distance from it.

```{r}
focal_vs_comp_ex %>% 
  tidyr::unnest(cols = "comp")
```

@@

Since the resulting output is an S3 object of class `comp_bayes_lm` we defined, we can both print and visualize its output: the posterior distribution of all linear regression parameters: the intercept, the slope for dbh for each species, and a matrix of all species pairs competitive effects on growth.

@@

Furthermore, we can apply a generic `predict()` function to the resulting `comp_bayes_lm` object to obtain fitted/predicted values of this model. We append these `growth_hat` values to our `focal_vs_comp` data frame.

```{r}
focal_vs_comp_ex <- focal_vs_comp_ex %>%
  mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))
focal_vs_comp_ex
```
We then compute the root mean squared error (RMSE) of the observed versus fitted growths as a measure of our model's fit.

@@

Whereas in our example above we fit our model to the entirety of the data and then generated fitted/predicted growths on this same data, we now apply the same model with spatial cross-validation. All the trees in a given fold will be given a turn as the "test" data while the trees is all remaining folds will be the "training" data. We then fit the model to the training data, but compute fitted/predicted growths for the separate and independent data.

>>>>

This package is designed to work for spatially mapped, repeat censused forests plots. The package has commands to fit models of tree growth based on neighborhood competition which can be used to estimate species-specific competition coefficients. The model fits can then be evaluated using a spatial cross-validation scheme to detect possible overfitting. Additionally, these models can test whether the species identity of competitors matters using a permutation test-style shuffling of competitor identity (under the null hypothesis) and subsequently evaluating if model performance changes. See Allen and Kim (2020) [A permutation test and spatial cross-validation approach to assess models of interspecific competition between trees](https://doi.org/10.1371/journal.pone.0229930) for a full description.

@@

```{r}
library(tidyverse)
library(forestecology)
library(sf)
library(sfheaders)
library(blockCV)
library(yardstick)
library(snakecase)
library(patchwork)
library(tidyr)
```

@@

The starting point of our analysis are data from two repeat censuses `census_1_ex` and `census_2_ex`. For example, consider the forest census data in `census_1_ex`.

@@

In the visualization below, the spatial blocks that act as our cross-validation folds are delineated in orange. The shape of each point indicates which fold each tree has been assigned to.

@@

Based on our `growth` data frame, we now explicitly define all "focal" trees and their respective "competitor" trees in a `focal_vs_comp` data frame. This data frame has rows corresponding to each focal tree, and all information about its competitors are saved in the list-column variable `comp`. We implemented this nested format using `nest()` in order to minimize redundancy, given that the same tree can act as a competitor multiple times.

@@

Using `unnest()` we can fully expand the competitor information saved in the `focal_vs_comp` data frame. For example, the tree with `focal_ID` equal to 2 located at (1.5, 2.5) has two competitors within `comp_dist` distance from it.

```{r}
focal_vs_comp_ex %>% 
  unnest(cols = "comp")
```

@@

The resulting output is an `comp_bayes_lm` object containing the posterior distribution of all linear regression parameters, the intercept, the slope for dbh for each species, and a matrix of all species pairs competitive effects on growth. The S3 object class is associated with several methods.

@@

Furthermore, we can apply a `predict()` method to the resulting `comp_bayes_lm` object to obtain fitted/predicted values of this model. We append these `growth_hat` values to our `focal_vs_comp` data frame.

```{r}
focal_vs_comp_ex <- focal_vs_comp_ex %>%
  mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))
focal_vs_comp_ex
```

We then compute the root mean squared error (RMSE) of the observed versus fitted growths as a measure of our model's fit.

@@

Whereas in our example above we fit our model to the entirety of the data and then generate fitted/predicted growths on this same data, we now apply the same model with spatial cross-validation. All the trees in a given fold will be given a turn as the "test" data while the trees in all remaining folds will be the "training" data. We then fit the model to the training data, but compute fitted/predicted growths for the separate and independent data.
